### 2023-10-02 doc/_intro.qmd
include Madhav's comments

We present a Bayesian inference pipeline to properly quantify the uncertainty
about the parametric information we can extract from high-throughput competitive
fitness assays. In these assays, the fitness of an ensemble of genotypes is
determined relative to a reference genotype [@kinsler2020; @ascensao2023].
@fig-01(A) shows a schematic of the experimental procedure in which an initial
pool of barcoded strains are mixed with a reference strain and inoculated into
fresh media. After some time---usually, enough time for the culture to
saturate---an aliquot is transferred to fresh media, while the remaining culture
is used for DNA sequencing of the lineage barcodes. The time-series information
of the relative abundance of each lineage, i.e., the barcode frequency depicted
in @fig-01(B), is used to infer the relative fitness---the growth advantage on a
per-cycle basis---for each lineage with respect to the reference strain. The
proposed statistical model accounts for multiple sources of uncertainty when
inferring the lineages' relative fitness values (see @sec-experiment for details
on sources of uncertainty accounted for by the model). Furthermore, minor
changes to the core statistical model allow us to account for relevant
experimental variations of these competition assays. More specifically, in
@sec-multienv, we present a variation of the statistical model to infer fitness
on growth dilution cycles in multiple environments with proper error
propagation. Furthermore, as described in @sec-replicates, our statistical model
can account for batch-to-batch differences when jointly analyzing multiple
experimental replicates using a Bayesian hierarchical model. Finally, a variant
of these hierarchical models, presented in @sec-genotypes, can account for
variability within multiple barcodes mapping to equivalent genotypes within the
same experiment.

For all the model variations presented in this paper, we benchmark the ability
of our pipeline to infer relative fitness parameters against synthetic data with
known ground truth. A `Julia` package accompanies the present method to readily
implement the inference pipeline with state-of-the-art scientific computing
software.

>>>>

We present a Bayesian inference pipeline to quantify the uncertainty about the
parametric information we can extract from high-throughput competitive fitness
assays given a model of the data generation process and experimental data. In
these assays, the fitness of an ensemble of genotypes is determined relative to
a reference genotype [@kinsler2020; @ascensao2023]. @fig-01(A) shows a schematic
of the experimental procedure in which an initial pool of barcoded strains are
mixed with a reference strain and inoculated into fresh media. After some
time---usually, enough time for the culture to saturate---an aliquot is
transferred to fresh media, while the remaining culture is used for DNA
sequencing of the lineage barcodes. The time-series information of the relative
abundance of each lineage, i.e., the barcode frequency depicted in @fig-01(B),
is used to infer the relative fitness---the growth advantage on a per-cycle
basis---for each lineage with respect to the reference strain. The proposed
statistical model accounts for multiple sources of uncertainty when inferring
the lineages' relative fitness values (see @sec-experiment for details on
sources of uncertainty accounted for by the model). Furthermore, minor changes
to the core statistical model allow us to account for relevant experimental
variations of these competition assays. More specifically, in @sec-multienv, we
present a variation of the statistical model to infer fitness on growth dilution
cycles in multiple environments with proper error propagation. Furthermore, as
described in @sec-replicates, our statistical model can account for
batch-to-batch differences when jointly analyzing multiple experimental
replicates using a Bayesian hierarchical model. Finally, a variant of these
hierarchical models, presented in @sec-genotypes, can account for variability
within multiple barcodes mapping to equivalent genotypes within the same
experiment.

For all the model variations presented in this paper, we benchmark the ability
of our pipeline to infer relative fitness parameters against synthetic data
generated from logistic growth simulations with added random noise. A `Julia`
package accompanies the present method to readily implement the inference
pipeline with state-of-the-art scientific computing software.
